Online {\IterComp}

### 2017-09-16 results.tex
Formatting
Results

The overhead for \OptProf typically remains below or around 10\% but can reach up to 59\% for \texttt{adpcm\_d}, which would be
unacceptable. \WCRelax-\textit{5}\% on the other hand manages to eliminate most of the overhead, with a worst
case of 16\% and an average of 8\%. \WPRelax-\textit{5}\% further reduces the cost of profiling, 5\% overhead on average and 13\%
worst case. For both \WCRelax and \WPRelax setting the threshold to 2\% makes it too difficult to remove probes. As a result, overhead
remain high.

@@

The benchmark where \OptProf records its worst result, \texttt{adpcm\_d}, has a single hot function which consists mainly of a single hot
loop with several branches. This leads to numerous short basic blocks, many of them with probes. Most of the improvement achieved by both
relaxation strategies comes from removing only two probes from the hot loop. They were initially placed in branches with a high probability
of being taken but with a small work contribution. Removing either of the probes introduces error of about 1.3\% at the loop level but
eliminates three quarters of the overhead.

For a few of the benchmarks, relaxation unexpectedly introduces additional overhead. This is counter-intuitive since relaxation leads to
less computational work. In these cases, relaxation did not manage to remove high frequency probes, so we did not benefit significantly.
On the other, the probes that were removed affected how \texttt{-O3} optimized the surrounding code, in some cases blocking useful
optimizations that were applied originally. This resulted in slower code.
%For example, some of the optimizations that make use of this built-in cost model
%and also alter the CFG are: loop-unrolling, function inlining, and
%simplifications of the CFG.
%These changes in the program's CFGs may affect other compiler optimizations and
%also possibly in hardware performance, for example affecting the instruction
%caching.

In Figure~\ref{fig:error-O3}, we see the error in our work estimation introduced by relaxation. The error for the WC-Relax strategies is
too low, always below 1\% and usually much lower, regardless of threshold. This confirms that WC-Relax is too conservative. WP-Relax brings
us closer to our target threshold. Despite not having hard guarantees for the error under WP-Relax, in all cases the error was below the
threshold.

@@

In this section we evaluate the effectiveness of our work efficiency metric.
We perform online {\itercomp} using all five profiling strategies, namely
\OracleRM, \OraclePP,\OptProf,\WCRelax-5\%, and \WPRelax-5\%.
Because the \OracleRM measures actual speedup, it provides reference results for
assessing the effectiveness of the work-efficiency metric.
For the relaxation strategies, we evaluate only those with a 5\% threshold, since
they offered better overhead reductions for little dynamic errors in the measurement.

@@

The comparison between \OracleRM and \OraclePP is crucial for validating the
effectiveness of the work efficiency metric in guiding online {\itercomp},
while the other configurations demonstrate the viability of applying online
{\itercomp} in real-world scenarios.

@@

The \OptProf achieves 4\% improvement on average, which represents 45\% of the
performance improvement obtained by the \OracleRM.
This difference to \OraclePP is explained by the real (intrusive) profiling used
in the \OptProf configuration.
The profiling affects the search in two key ways:
\textit{(i.)} the overhead incurred by the profiling affects the execution time,
and therefore also affects the work efficiency metric,
\textit{(ii.)} the instrumentation code may affect many of the optimizations, e.g.,
due to the use of a global variable or in decisions taken based on a cost-model
for the instructions.
However, even in the \OptProf configuration, which has a highly intrusive instrumentation,
it is still possible to obtain performance improvements in a realistic online
{\itercomp}.

Finally, the evaluation also indicates that the use of relaxation algorithms
is beneficial not only for its reduction in overhead, which is directly observed
by the user, by it also tends to improve the quality of the search.
This improvement has to do with the aforementioned effects of an intrusive
instrumentation on the search.
The relaxation has the potential to reduce both effects of the profiling.
Both the \WCRelax-5\% and \WPRelax-5\% achieve about 60\% of the
performance improvement obtained by the \OracleRM.

>>>>

The overhead for \OptProf typically remains below or around 10\% but can reach up to 59\% for \texttt{adpcm\_d}, which would be
unacceptable. \WCRelax-\textit{5}\% on the other hand manages to eliminate most of the overhead, with a worst
case of 16\% and an average of 8\%. \WPRelax-\textit{5}\% further reduces the cost of profiling, 5\% overhead on average and 13\%
worst case. For both \WCRelax and \WPRelax setting the threshold to 2\% makes it too difficult to remove probes. As a result, overhead
remains high.

@@

The benchmark where \OptProf records its worst result, \texttt{adpcm\_d}, has a single hot function which consists mainly of a single hot
loop with several branches. This leads to numerous short basic blocks, many of them with probes. Most of the improvement achieved by both
relaxation strategies comes from removing only two probes from the hot loop. They were initially placed in branches with high probabilities
of being taken but with a small work contribution. Removing either of the probes introduces error of about 1.3\% at the loop level but
eliminates three quarters of the overhead.

For a few of the benchmarks, relaxation unexpectedly introduces additional overhead. This is counter-intuitive since relaxation leads to
less computational work. In these cases, relaxation did not manage to remove high frequency probes, so we did not benefit significantly.
On the other, the probes that were removed affected how \texttt{-O3} optimized the surrounding code, in some cases blocking useful
optimizations that were applied originally. This resulted in slightly slower code.
%For example, some of the optimizations that make use of this built-in cost model
%and also alter the CFG are: loop-unrolling, function inlining, and
%simplifications of the CFG.
%These changes in the program's CFGs may affect other compiler optimizations and
%also possibly in hardware performance, for example affecting the instruction
%caching.

In Figure~\ref{fig:error-O3}, we see the error in our work estimation introduced by relaxation. The error for the WC-Relax strategies is low, always below 1\% and usually much lower, regardless of threshold. This confirms that WC-Relax is very conservative. WP-Relax brings
us closer to our target threshold. Despite not having hard guarantees for the error under WP-Relax, in all cases the error was below the
threshold.

@@

In this section we evaluate the effectiveness of our work efficiency metric.
We perform online {\itercomp} using all five profiling strategies, namely
\OracleRM, \OraclePP, \OptProf, \WCRelax-5\%, and \WPRelax-5\%.
Because the \OracleRM measures actual speedup, it provides reference results for
assessing the effectiveness of the work-efficiency metric.
For the relaxation strategies, we evaluate only those with a 5\% threshold, since
they offered better overhead reductions for little dynamic error in the measurement.

@@

The comparison between \OracleRM and \OraclePP is crucial for validating the
effectiveness of the work efficiency metric in guiding online {\itercomp},
while the other configurations demonstrate the viability of applying online
{\itercomp} in real-world scenarios. \OracleRM manages an average speedup of 9.3\% over \texttt{-O3}. This is modest by typical \itercomp
standards because search is a relatively small random search. The search driver is not the focus of this work, and we expect higher speedups from more complex drivers, like genetic algorithms.

@@

The \OptProf achieves 4\% improvement on average, which represents 45\% of the
performance improvement obtained by the \OracleRM.
This difference to \OraclePP is explained by the real (intrusive) profiling used
in the \OptProf configuration.
The profiling affects the search in two key ways:
\textit{(i.)} the overhead incurred by the profiling affects the execution time,
and therefore also affects the work efficiency metric,
\textit{(ii.)} the instrumentation code may affect many of the optimizations, e.g.,
due to the use of a global variable or in decisions taken based on a cost-model
for the instructions.
However, even in the \OptProf configuration
it is still possible to obtain performance improvements in a realistic online
{\itercomp}.

Finally, the evaluation also indicates that the use of relaxation algorithms
is beneficial not only for its reduction in overhead, which is directly observed
by the user, by it also tends to improve the quality of the search.
This improvement has to do with the effects of 
instrumentation on the search mentioned earlier.
The relaxation has the potential to reduce both effects of the profiling.
Both the \WCRelax-5\% and \WPRelax-5\% achieve about 60\% of the
performance improvement obtained by the \OracleRM.
